---
title: "chapter6.rmd"
author: "Nagabhooshan Hegde"
date: "03/12/2020"
output: html_document
---
Chapter 6

Task 1: Analysis of RATS data

```{r}

library(stats)
library(dplyr)
library(tidyr)
library(ggplot2)
```


Let us load RATS data and look at its structure

```{r}
url_RATS="https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt"

RATS=read.csv(url_RATS, sep="\t")
RATS$ID=as.factor(RATS$ID)
RATS$Group=as.factor(RATS$Group)
glimpse(RATS)

```


Get the data into long form

```{r}
RATSL <- RATS %>%
  gather(key = WD, value = Weight, -ID, -Group) %>%
  mutate(Time = as.integer(substr(WD,3,4))) 

str(RATSL)
```


Let's plot the data

```{r}
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)), name = "Weight (grams)")+
  scale_x_continuous(name = "Time (days)")
```


Plot shows Groups have significant difference. Group 1 is seperated from 2 and 3. Group 2 has a clear outlier, and there is not much difference between Group 2 and 3

Now let's try to standardize and plot the data

```{r}
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate( stdWeight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()
glimpse(RATSL)
```


```{r}
ggplot(RATSL, aes(x = Time, y = stdWeight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(name = "Standardized Weight (grams)")+
  scale_x_continuous(name = "Time (days)")+
  ggtitle("Standardized weight of rats in three different diet groups")
```


Let us plot boxplot of the data

```{r}
ggplot(RATSL, aes(x = factor(Time), y = Weight, fill = Group)) +
  geom_boxplot(position = position_dodge(width = 0.9)) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(legend.position = c(0.9,0.4))+ 
  scale_x_discrete(name = "week") +
  scale_fill_grey(start = 0.5, end = 1)
```


This plot shows clear difference between groups and weight is gradually increasing over time. There are many outliers in each groups

Lets calculate mean for all groups

```{r}
RATS2 <- RATSL %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

glimpse(RATS2)
```

Plot boxplot of the mean Vs group

```{r}
ggplot(RATS2, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(weight)")
```

Most clear outlier group is 2, so let's remove it

```{r}
#Create new data removing outlier
RATS3 <- RATS2 %>%
  filter(mean < 550)

ggplot(RATS3, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(weight), over time period")


```


This shows standard deviation within group is small but differences between groups can be seen clearly


```{r}
fit <- lm(mean ~ Group, data = RATS3)
summary(fit)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit)

```


After examining this result, we can see that there are statistically significant differences between the diet groups




Task 2 : Analysis of BPRS data

```{r}
url="https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt"
BPRS=read.csv(url, sep=" ")

BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)

# Convert to long form
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)

# Extract the week number
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))
glimpse(BPRSL)

```


Lets plot the data

```{r}
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject))+ 
  geom_line() + 
  scale_linetype_manual(values = rep(1:10, times=4)) + 
  facet_grid(. ~ treatment, labeller = label_both) + 
  theme_bw() + 
  theme(legend.position = "none") + 
  theme(panel.grid.minor.y = element_blank()) + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```


```{r}
#Create regression model 
BPRSL_reg <- lm(bprs ~ week + treatment, data = BPRSL)
summary(BPRSL_reg)
```


This shows week variable is statistically significant

```{r}
library("lme4")
# Creating a random intercept model with subject as random factor
BPRSL_ref1 <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRSL_ref1)

```


Summary shows standard deviation is high that means there is variablity between subjects

```{r}
BPRSL_ref2 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)
summary(BPRSL_ref2)

```


performing an ANOVA test on the two models

```{r}

anova(BPRSL_ref2, BPRSL_ref1)
```



Create a vector of the fitted values

```{r}
Fitted <- fitted(BPRSL_ref2)
BPRSL <- BPRSL %>%
  mutate(Fitted)
```


```{r}
ggplot(BPRSL, aes(x = week, y = Fitted, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))

```

So this shows both original and fitted values decrease over week