---
title: "chapter3"
author: "Nagabhooshan Hegde"
date: "12/11/2020"
output: html_document
---

```{r}
url = "http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt" 
alc=read.csv(url, sep=",")
glimpse(alc)
```
Data consists student information from 2 different schools.Data was collected by school reports and quetionnaires.This data has numerical, categorical and logical values.
Data has 382 rows and 35 columns
4 variables to consider: famrel , freetime ,goout ,absences
Family relationship:students with high value may consume less alcohol
freetime: more free time more going out and consuming alcohol
goout: more going out and more consuming alcohol
absences :more absences more alcohol consumption

Let's draw bar chart of all variables, and box plots of variables we are considering
```{r}
gather(alc) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
#boxplot for family relationship
g1 = ggplot(alc, aes(x = high_use, y = famrel))
g1 = g1 + ggtitle("Student family relation vs alcohol consumption")
g1 + geom_boxplot() + ylab("family relation")

#box plot for freetime
#par(mfrow=c(2,2))
g2 = ggplot(alc, aes(x = high_use, y = freetime))
g2 = g2 + ggtitle("Student freetime vs alcohol consumption")
g2 + geom_boxplot() + ylab("freetime")

#boxplot for students going out 
g3 = ggplot(alc, aes(x = high_use, y = goout))
g3 = g3 + ggtitle("Student go out vs alcohol consumption")
g3 + geom_boxplot() + ylab("go out")

#boxplot for absence of student
g4 = ggplot(alc, aes(x = high_use, y = absences))
g4 = g4 + ggtitle("absence of student vs alcohol consumption")
g4 + geom_boxplot() + ylab("absence")

```
After looking at boxplot, it is clear that free time of student has no effect on alcohol consumption. where as other factors shows relation with alcohol consumption
```{r}
summary(m)
# print out the coefficients of the model
coef(m)
# compute odds ratios (OR)
OR <- coef(m) %>% exp

# compute confidence intervals (CI)
CI <- confint(m) %>% exp

# print out the odds ratios with their confidence intervals
cbind(OR, CI)

```
Family relation has negative correlation.
Freetime has no effect 
Go out has positive correlation
Absence has positive correlation but very minimal
Summary shows free time has no effect. and other factors have some effects they can be good parameters
```{r}
# predict() the probability of high_use
probabilities <- predict(m, type = "response")
# add the predicted probabilities to 'alc'
alc <- mutate(alc, probability = probabilities)

# use the probabilities to make a prediction of high_use
alc <- mutate(alc, prediction = probability > 0.5)

# see the last ten original classes, predicted probabilities, and class predictions
select(alc, failures, absences, sex, high_use, probability, prediction) %>% tail(10)

# tabulate the target variable versus the predictions
table(high_use = alc$high_use, prediction = alc$prediction)
```