---
title: "chapter5"
author: "Nagabhooshan Hegde"
date: "26/11/2020"
output: html_document
---

This exercise is based on human data which consists of Human Development Index. More information of the data can be found here
```{r}
library(corrplot)
library(tidyverse)
library(ggplot2)
library(FactoMineR)
library(GGally)
url="http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt"
human=read.csv(url)
summary(human)

```


```{r}
library(ggplot2)
ggpairs(human)
```


Now this shows distributions and relationship between variables.
Expected year of schooling is normally distributed
Distribution of some of the variables are on left Ex. : GNI, Mat-Mor,Ado.Birth and Parli.F
Distribution of some of the variables are on right Ex. : Labo.FM, Life.EXP
Summary below shows mean, median, min and max values of the variables
```{r}
summary(human)
```


Compute the correlation matrix and visualize it with corrplot:

```{r}
cor(human) %>% corrplot(type = "upper", tl.cex=0.6)
```


variables ‘Labo.FM' and 'Parli.F' don't have significant correlation with other variables
Life.Exp and Mat.Mor have highest correlation
```{r}
cor(human)

```


PCA(Principal Component Analysis)

Perform principal component analysis (PCA) on the not standardized human data
```{r}
pca_human <- prcomp(human)
s <- summary(pca_human)

# rounded percetanges of variance captured by each PC
pca_pr <- round(1*s$importance[2, ]*100, digits = 1)

# create object pc_lab to be used as axis labels
pc_lab=paste0(names(pca_pr), " (", pca_pr, "%)")

# draw a biplot
biplot(pca_human, cex = c(0.5, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])

```


it clearly shows that GNI is only explaining variable for non-standardized data.

Standardize the variables in the human data

```{r}
human_std <- scale(human)

## PCA model and biplot
pca_human <- prcomp(human_std)

s1 <- summary(pca_human)

# rounded percetanges of variance captured by each PC
pca_pr <- round(1*s1$importance[2, ]*100, digits = 1)

# create object pc_lab to be used as axis labels
pc_lab=paste0(names(pca_pr), " (", pca_pr, "%)")

# draw a biplot
biplot(pca_human, cex = c(0.5, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])
```


Comparision between standardised and non standardised PCA
PCA with non standardised data has only one arrow of GNI, but PCA with standardised data shows multiple arrows pointing at different directions
In PCA of standardised data, PCA1 explains 53.6% of variance and PCA2 explains 16.2% of variance

MCA

```{r}
data(tea)

print(str(tea))
```

for simplicity lets consider only few variables

```{r}
keep_columns <- c("Tea", "How", "sex", "sugar","where", "lunch")
tea_time <- subset(tea, select=keep_columns)
# visualize the dataset
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
print(summary(tea_time))
```

multiple correspondence analysis

```{r}
#MCA
mca <- MCA(tea_time, graph = FALSE)
# visualize MCA
plot(mca, invisible=c("ind"))
```


Multiple Correspondence Analysis (MCA) is a method that allows studying the association between two or more qualitative variables

No sugar is close to females

sugar is close to males 
